---
title: "P8105 Homework 6"
output: github_document
date: "November 23, 2024"
---

```{r document setup, include = FALSE }
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)
library(tidyverse)
library(modelr)
set.seed(1000)
```

## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Perform 5000 bootstrap samples and fit linear models
```{r}
bootstrap_results <- weather_df %>%
  modelr::bootstrap(n = 5000, id = "bootstrap_id") %>%
  mutate(
    models = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    # Extract R-squared
    r_squared = map_dbl(models, ~ glance(.x)$r.squared),
    # Compute log(beta0 * beta1)
    log_beta_product = map_dbl(models, ~ {
      coefs <- tidy(.x)$estimate
      log(coefs[1] * coefs[2])
    })
  )
```

Compute confidence intervals
```{r}
ci_results <- bootstrap_results %>%
  summarise(
    r_squared_lower = quantile(r_squared, 0.025),
    r_squared_upper = quantile(r_squared, 0.975),
    log_beta_lower = quantile(log_beta_product, 0.025),
    log_beta_upper = quantile(log_beta_product, 0.975)
  )
```
The 95% confidence interval for r squared is `r ci_results$r_squared_lower`, `r (ci_results$r_squared_upper)`.
The 95% confidence interval for log(beta0*beta1) is `r ci_results$log_beta_lower`, `r ci_results$log_beta_upper`


Plot distributions
R-squared distribution
```{r}
r_squared_plot <- ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(binwidth = 0.005, fill = "blue", alpha = 0.7) +
  labs(
    title = "Bootstrap Distribution of R-squared",
    x = expression(R^2),
    y = "Frequency"
  ) +
  theme_minimal()
```

Log(beta0 * beta1) distribution
```{r}
log_beta_plot <- ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram(binwidth = 0.05, fill = "yellow", alpha = 0.7) +
  labs(
    title = "Bootstrap Distribution of log(beta0 * beta1)",
    x = expression(log(beta[0] * beta[1])),
    y = "Frequency"
  ) +
  theme_minimal()
```

Display the plots
```{r}
print(r_squared_plot)
print(log_beta_plot)
```

The bootstrap distribution of R² is slightly skewed to the right, with most values falling between 0.89 and 0.93 and a mode near 0.91, suggesting a strong and consistent model fit across different samples. The distribution of log(β₀β₁) is approximately symmetric and normally distributed, centered around 2.00, indicating stable relationships between the intercept and slope across re-samples.



